---
title: "TBtypeR"
output: 'github_document'
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, results = "hold", message = FALSE, warning = FALSE)
```

`TBtypeR` is an R package that performs accurate and sensitive
quantification of mixtures of *M. tuberculosis* (MTB) strains from whole
genome sequencing (WGS) data. `TBtypeR` excels as detecting low-frequency mixed infections that other tools struggle to detect, maintaining a sensitivity of 85% down to minor strain frequencies of 2.5%. 

## Running TBtypeNF
`TBtypeNF` is an end-to-end [Nextflow](https://www.nextflow.io/index.html) pipeline for `TBtypeR` that takes FASTQ  files as input and performs FASTQ preprocessing with [fastp](https://github.com/OpenGene/fastp), read alignment with [BWA-MEM](https://github.com/lh3/bwa), variant calling with [BCFtools](https://samtools.github.io/bcftools/bcftools.html), and quality control (QC) report generation using [SAMtools](https://www.htslib.org/), [mosdepth](https://github.com/brentp/mosdepth). The output is an HTML report with detected MTBC strains and mixtures frequencies.

`TBtypeNF` requires a sample manifest to run, minimally a TSV file with column names "sample", "fastq1" and "fastq2" - see [example manifest](TBtypeNF/resources/lung_example_manifest.tsv).

### Requirements  
* Nextflow (≥ 22.03.0)
* Singularity (Apptainer) or Docker

### Usage  
```{bash, eval=FALSE}
# download example manifest
wget https://raw.githubusercontent.com/bahlolab/TBtypeR/main/TBtypeNF/resources/lung_example_manifest.tsv -O my_manifest.tsv
# run the nextflow pipeline
nextflow run bahlolab/TBtypeR/TBtypeNF/main.nf -r main -profile singularity --manifest my_manifest.tsv
```

## Using TBtypeR
The recommended way to use `TBtypeR` is through the `TBtypeNF` pipeline. However, additional parameters and customisation is available by using the R package directly. `TBtypeR` can be installed with `devtools` as follows:
```{r, eval=FALSE}
devtools::install_github("bahlolab/TBtypeR")
```

Example usage of `TBtypeR` as follows:
```{r detect_mixtures}
library(SeqArray)
library(tidyverse)
library(TBtypeR)

# replace with path to your VCF file
vcf_filename <- system.file('vcf/example.vcf.gz', package = 'TBtypeR')

tbtype_result <- 
  tbtype(vcf = vcf_filename) %>% 
  filter_tbtype(max_phylotypes = 3) %>%
  unnest_mixtures()

tbtype_result %>% 
  knitr::kable()
```
```{r visualise_result}
tbtype_result %>% 
  ggplot(aes(x = sample_id, y = mix_prop, fill = mix_phylotype)) +
  geom_col() +
  coord_flip()
```

